Considerations for the Next Iteration of the DOE Fast Forward Storage and IO Project

### 2014-02-28 paper.tex
relatively minor edits. Some more notes of sections that need help. The others are in pretty good shape.

\begin{abstract}
The Fast Forward effort to define a next generation filesystem has made
admirable contributions in architecture and design. Formalizing the general
idea of data staging as burst buffers for the file system will help manage
the performance variability and offer the data processing opportunities
outside the main compute and file system. Adding a tranasctional mechanism to
manage faults and data visilibilty helps enable effective analytics without
having to work around the IO stack semantics.

@@

This paper examines some of the choices made by the FastForward team and
compares them with other options and offers observations and suggestions based
on these other efforts.  This will include some non-core contributions, such
as some of the demonstration components generated while implementing \DDT,
to make suggestios that may help the next generation design for how the IO
stack works as a whole.

@@

Current production HPC IO stack design is unlikely to offer sufficient features
and performance to adequately serve the needs of an extreme scale platform. To
address these limitations, the US Department of Energy commissioned an effort
to develop a design and prototype for an IO stack suitable for the extreme
scale environment. This is a joint effort led by the Intel Lustre team, Los
Alamos National Laboratory, EMC, and the HDF Group. This team has developed a
specification set~\cite{fastforward:2014:docs} for a future IO stack to address
the identified challenges.

@@

The core idea for IOD is to provide a way to manage the IO load that is
separate from the compute nodes and the storage array. Communication intensive
acitivies, such as two-phase, data sieving IO's data rearrangement process, can
be moved to the IOD layer offloading the communication load from the compute
nodes. IOD has three main purposes. First, if the optional burst buffer is
availble, it works as a fast cache absorbing write write operations for the
slower trickle out to the central storage array. It can also be used to
retrieve objects from the central storage array for more efficient read
operations and offer data filtering to make client reads more efficient.
Second, it offers the transaction mechanism for controlling data set visibility
and to manage faults. Third, data processing operations can be placed in the
IOD. This operations are inteded to offer data rearrangement, filtering, and
similar operations prior to it reaching the central storage array.

This offloads the requirement for collective two-phase data sieving
on the compute nodes to reorganize data. This technique has been proven
effective at reducing the total time for writing data due to fewer participants
involved in the communication patterns~\cite{lofstead:2011:nessie-staging}.
Beyond these broad items, there are many important details some of which are
examined in more detail below.

The bottom two layers are the Data Access Object Storage (DAOS) and Virtual
Object Storage Device (VOSD). DAOS is the typical interface that represents
the individual objects stored including the semantics related to ``files''.
VSOD is the actual storage mechanism for the objects defined at the DAOS layer.

The focus of this paper is primarily the IOD layer given the critical role it
has in the performance and functionality of the entire stack. Some of the key
features explored in this paper all have a focus on the IOD layer motivating
the focus of this examination.

Incorporated into the design of these layers and the function of some of these
layers themsleves, such as the IOD layer, are features to manage both burst IO
performance and faults. In particular, the IOD layer's use of burst buffers
attempts to absorb a massive data dump from the compute area that is then
trickled to the persistent storage area only if desired. Otherwise, it can be
stored locally on optional SSDs for use by analysis routines eliminating the
need to use slower rotational media as the intermediary between a simulation
and the analysis code.

A second feature is the incorporation of a transaction like mechanism into the
IOD layer and a related epochs feature into the DAOS layer. This feature offers
a mechanism to avoid making incomplete data visible to other users and offer a
level of protection if the data can be stored on the optional SSDs. The
implications of how these concepts are designed are important to consider for
the final design. Because they control data visibility and manage faults for
output operations, they are key features.

The rest of the paper is organized as follows. Section~\ref{sec:burst}
discusses some of the features of incoporating burst buffers as designed and
suggests some considerations and alternatives for the next generation of this
project. Section~\ref{sec:transactions} discusses the transactions approach
offered in the IOD layer and the corresponding epochs in the DAOS layer. It
also offers a comparison to the \DDT system given the very similar high-level
design and motivating use case.  Section~\ref{sec:optimizations} discusses some
of the optimizations offered by the current design and how and when these
optimizations will work best. It also explores some of the dark corners that
may yield performance issues with the current design. Finally,
Section~\ref{sec:summary} disucsses the system overall with recommendations on
what design elements should be considered based on broader issues with current
HPC data centers.

@@

By concentrating this functionality into the storage stack, two problems arise.
First, the amount of network bandwidth, IO bandwidth, and compute power
consumed for example operations from a single application is likely to
completely monopolize the IOD processes. Second, if space and time partitioning
is used instead, the functionality risks being too small to be useful. Third,
the hardware performance or SSDs is questionable. Current NAND-based flash
devices top out at around 400 MB/sec. The key spec that is missing from this
number is that 400 MB/sec is a measure of the fixed number of available IOPS
multipled by the block size. This represents the ideal streaming performance
possible. The problem is that it costs an IOP to read 1 byte or 1 block. It
costs 1 IOP to write a full block--usually. In some cases, it will cost 2 IOPs.
This accounts for the required pre-erase write prior to writing the block. In
the worst case, it can be 3 IOPS per write. This would be for a partial block
write (read the old block, erase, write the new block). One-third of 400
MB/sec, about 133 MB/sec, is well below the streaming performance of HDDs.
Granted, there is still rotational and seek latency to deal with for HDDs, but
the advantage for SSDs has evaporated and potentially turned into a penalty.

Given these features, the optionality and even incorporation of burst buffers
in the current design should be reconsidered. Much of the advanced, key
functionality proposed as they are currently designed ultimatey relies on the
existence of burst buffers to work. Further thought about how to have an IOD
layer both with and without a burst buffer is required before they can be
considered optional. As the design stands today, they are a required part of
the IOD layer for proper functioning. Unfortunately, it is not clear that they
can address the performance concerns they are intended to cover.

@@

The transaction mechanism manifests in two forms. At the IOD layer, they are
called transactions and are used to judge whether or not an output is complete
or not and control access through treating the transaction ID as a version
identifier. At the DAOS layer, they are called epochs and represent persisted
(durable) transactions from the IOD layer. Each of these offers different
functionality, but are connected as is explained below. How these differ from
the \DDT approach is also explored. While they are seemingly very different,
they use a very similar approach to solve the same problem.

@@

\subsection{DAOS Epochs}
The Epoch mechanism differs from transactions. Instead of focusing on when a
particular output is complete, an epoch represents incremental persisted copies
of a container. To simplify the mapping between an IOD transaction and the DAOS
epochs, the IOD transaction ID is the same as the epoch ID. The key difference
is that at the DAOS layer, some transaction (epoch) IDs will not be
represented.

@@

While a pure no metadata model would be ideal for performance, this current
proposal has a few challenges. By providing this basic functionality, metadata
management is split between the user layer and the IOD layer. Consdier the
following items. First, there are no list of containers. This is closer to the
no metadata ideal, but given the rest of the functionality, it exports to the
user level a requirement to track what containers are in the file system.
Second, given the flattening that occurs between IOD and DAOS, the mapping from
IOD objects to DAOS objects is unlikely to be 1-to-1. This is likely a list
instead. Further, given the ability to re-stripe in the IOD layer, the list is
not a list of objects completely included in the IOD object, but instead
objects from which part of the data came from some listed DAOS object.  Third,
sharding IOD objects makes sense based on the compute process connection a
shard has with the complete object. Striping of an IOD object makes sense in
connection with the DAOS object(s) underneath, but striping within the IOD
layer is not discussed. Further, using striping from clients into the IOD layer
introduces a level of coordination that IOD seeks to avoid.  Fourth, given the
previous items, there is ambiguity with the maximum valid offset of the object.
This could be for the shard or for the collection of shard objects that
represent a distributed variable. Ideally, either a robust metadata management
system should be incorporated into the system, preferably separate from the IOD
layer. Alternatively, all of this functionality should be left to the user
layer. The only potential challenge is determining the container for a file
open command and what the transaction ID should be. This limited operation
should be a straightforward search using a hash or similar fast mechanism for
examining a list of container names and/or object IDs.

@@

@@

First, \DDT has a timeout mechanism to detect failures and offer an ability to
recover and clean up the incomplete transactions. IOD's transactions are
asynchronous precluding a relatively simple, short timeout value to detect
failures. Unfortunately, at this point, there is no mechanism in IOD's
transaction handling to detect a fault and clean up any incomplete operations.

@@

Third, in the single leader model, if the process that manages the transaction
were to fail, there is no ability to abort that transaction. The data will not
be made available because the transaction is still listed as in progress. The
problem is that without a way to clean up this failed transaction, no newer
versions of the same container can be made readable. In reality, the single
leader model would ideally outsource the transaction processing to something
like \DDT on the client side.

@@

\DDT has addressed this count issue in a couple of ways. First, the
sub-coordinators each have a list of processes from which they expect messages.
Should a message be missed, it is noticed and corrective action can be taken.
Second, \DDT has the concept of sub-transactions. The messaging requirements
are illustrated in Figure~\ref{fig:optimized-protocol}.  Sub-transactions
represent finer grained operations than the entire output, \DDT can manage
multiple writes per client by using a sub-transaction to represent the output
for any item to the file (container). Because of how the sub-transactions are
managed, the singleton sub-transactions must be declared before the transaction
begins ensuring there is global knowledge that the sub-transaction is expected.
That way if the coordinator (transaction leader) fails, which ever process
takes over that role knows to expect a completion message for that
sub-transaction or the overall transaction cannot complete. While this
additional layer does introduce messaging, the overhead is quite small.

@@

As an alternative, a scalable locking service like Chubby~\cite{burrows:2006:chubby} might be employed.

@@

@@

\section{Conclusions}
\label{lab:conclusion}

@@

The partial metadata management incorporated into the IOD layer and the lack of
consideration for how to handle and recover from failures are glaring
oversights in the current documents. It is our undertanding that these will be
addressed in the next phase and we hope to help inform that effort with our
experiences.

@@

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\begin{figure}[ht]
%\vspace{-0.15in}
%\centering
%\includegraphics[keepaspectratio=true, width=0.9\columnwidth]{images/hierarchy}
%\vspace{-0.15in}
%\caption{Example Hierarchy}
%\label{fig:hierarchy}
%\vspace{-0.15in}
%\end{figure}

%\begin{figure}[ht]
%\centering
%\includegraphics[keepaspectratio=true, width=0.9\columnwidth]{images/stagingmodel}
%\vspace{-0.15in}
%\caption{Staging Model}
%\label{fig:staging-model}
%\vspace{-0.15in}
%\end{figure}

%\section{New Protocol Design}
%\label{lab:design}

%To summarize the semantics of \DDT for those unfamiliar with it, consider the
%following. Since we have parallel clients and servers (MxN), it is necessary to
%have a mechanism to manage each single operation by a collection of clients 
%with a collection of servers to ensure that it is completed correctly. We must
%also have a way to manage the set of these operations as a single, atomic
%action. The difficulty with this scenario is the need to maintain global
%knowledge across all of the clients and servers. This global knowledge allows
%decisions across all clients. In a traditional distributed transaction, the
%single client always knows the state reported by all of the ``server''
%resources used. When the client is split into multiple pieces, each piece
%ultimately needs to have the same knowledge, including any knowledge that might
%be local only. If this does not happen, no guarantees can be made. The message
%count explosion implied by this level of knowledge is daunting and what
%prompted this project. Adding complexity for data movement operations is that
%a typical data set will contain multiple globally distributed variables that
%each must be treated as an atomic unit.

%To accommodate these requirements, we have the concept of a transaction
%representing the total set of operations and the sub-transaction representing a
%single operation. Each sub-transaction can represent either a single process
%performing an action (singleton) or all of the parallel clients participating
%in a single action (global). In either case, each sub-transaction is managed as
%part of the overall transaction. Voting and commit operations are performed
%both at the sub-transaction level and at the transaction level. With the
%re-implementation of the protocol, these semantics have not changed.

%\section{Evaluation}
%\label{lab:evaluation}

%To evaluate the scalability and performance of \DDTns, tests are performed on
%Cielo at Sandia National Laboratories. Cielo is a Cray XE6 platform with 8,944
%nodes each containing dual, eight core AMD Opteron 6136 processors and 32 GB of
%RAM. It uses the Cray Gemini network in a 3D torus and uses the Cray Linux
%Environment. Tests are performed at powers of two scale from 256 up to 65536
%processes. We use two pairs of datastore and metadata services along side the
%process counts mentioned here. Each of these other services is given a single
%node on which to operate.

%The represented IAW consists of a writing application creating ten 3-D
%variables. Each of these variables is distributed across all of the processes
%with each process containing a 32x32x32 double precision floating point number
%chunk for each variable. The total space is scaled by doubling each dimension
%keeping the total dimensions as close to square as is possible. All ten of
%these variables and the corresponding chunks are written to the first pair of
%datastore and metadata servers. Then, an update application marks one variable
%as ``in process'' in the first metadata service, reads the chunks in parallel
%from the first datastore service, updates the data, creates a new variable in a
%second metadata service inserting the updated data into the second datastore
%service, and then deletes the original variable from the first datastore and
%metadata services. For the purposes of this evaluation, only the overhead
%incurred by the transaction protocol is reported.  No attempts were made to
%make the datastore and metadata service efficient and including those
%performance results would obfuscate the impact of having transactions surround
%data movement applications. The overheads are measured during the update
%application's activities and represent the green and blue arrows within
%Figure~\ref{fig:optimized-protocol}.

%With the three-tier structure employed by \DDTns, we choose to always employ at
%least two sub-coordinators, even at small scale, to incur higher overhead
%costs.  An example of this configuration is shown in
%Figure~\ref{fig:hierarchy}.  In this case, process 0 acts as the coordinator,
%sub-coordinator, and a subordinate. Process 3 acts as a sub-coordinator and a
%subordinate. The rest of the processes operate solely as subordinates. The
%configuration for each run consists of a minimum of 2 sub-coordinators and a
%maximum of 256 subordinates per sub-coordinator. That yields 256
%sub-coordinators each with 256 subordinates at 65536 processes. This is
%illustrated in Table~\ref{tab:scaling} showing the various scales evaluated and
%the configuration for each.  Other tests of smaller numbers of processes were
%run, but were omitted because they did not add any additional information to
%the results.

%The test values reported include the following actions:

%\begin{enumerate}
%\item txn\_create\_transaction - create a new transaction
%\item txn\_create\_sub\_transaction - create a new singleton sub-transaction
%\item txn\_create\_sub\_transaction\_all - called three times to create three global sub-transactions
%\item txn\_begin\_transaction - start the transaction processing
%\item txn\_commit\_sub\_transaction - called four times total to commit the four sub-transactions
%\item txn\_vote\_transaction - vote to commit or abort the overall transaction
%\item txn\_commit\_transaction - commit the entire transaction
%\item txn\_finalize\_txn - clean up any transaction specific memory or perform any final operations
%\end{enumerate}

%All operations except txn\_create\_sub\_transaction and \\
%txn\_commit\_sub\_transaction are performed as effectively collective
%operations. These two exceptions are only performed by the single process
%involved. In order for the existence of this singleton sub-transaction to be
%guaranteed to be known, we split the create and begin operations to provide a
%timeframe during which all singleton sub-transactions should be initialized.
%The begin operation gathers knowledge of all of these singleton
%sub-transactions and broadcasts it to all processes to make failure recovery
%possible. This step guarantees that every singleton sub-transaction will be
%known or the begin will fail due to a process failure. Since the global
%sub-transactions are global operations, they can be done either before or after
%the begin transaction call. In our case, we have one global sub-transaction
%created before the begin transaction and two afterwards.

%The selected sequence of operations gives a representative idea of what a
%typical transaction may include. Each test is performed seven times and the
%mean value is presented. Given the small magnitude of the values, there is a
%bit of variance that has a large percentage, but small absolute value. The
%results are presented in Figure~\ref{fig:performance}. The results for 4096
%processes were lost.

%In the interest of saving space, only the results for 64K processes is
%discussed. In all cases, the time spent in operations for the 64K case are the
%longest of all cases tested. At a detailed level, the time for the 64K
%processes case can be itemized as follows. First, the slowest operation, by
%far, is the txn\_create\_sub\_transaction\_all with a maximum time of 0.0310
%seconds for one of the calls. The mean time is 0.01 seconds. Second, the other
%operations are all at less than 0.005 seconds on average and a maximum of 0.012
%seconds.  Third, the time for the initialization and cleanup of the total
%protocol, the calls equivalent to MPI\_Init and MPI\_Finalize, take 0.38
%seconds total. Since the txn\_init call is the first call that uses MPI in the
%system and that it is a full order of magnitude slower than any other operation
%at worst, we believe there is MPI initial communication channel setup being
%established accounting for the extra time. The txn\_finalize takes a maximum of
%0.0002 seconds.  Since these calls are done once for the entire application
%lifetime, we did not include them in the results since they are not part of the
%typical transaction cost that would be incurred for each transactional
%operation.

%\section {Discussion}
%\label{lab:discussion}
%The leading alternative to \DDT is the Intel/Lustre FastForward Epoch
%approach~\cite{barton:2013:fastforward,lombardi:2013:epochs}.  While \DDT is
%intended for a general purposes, epochs are designed specifically to address
%the needs of a shared storage array with a large number of parallel clients.
%The general idea is to mark every operation with a transaction id, record a
%process count somewhere to allow reconciliation later, and have every process
%operate independently for the duration of the writing operation. A second, more
%synchronous mode is also available. In this synchronous mode, the user is
%required to manage the transaction ID and to update the transaction state in
%the storage servers. During a read process, it must be determined what is the
%most current, complete, ``Readable'' version of the file.  This entails
%contacting all of the storage devices to determine all of the versions of
%blocks stored on each device. With that information, it can be determined which
%of these transactions is ``Readable''. While this reconciliation operation may
%happen asynchronously, it must be done before a read operation potentially
%prompting it to be triggered by a client call or potentially stale data is
%presented as the most current.

>>>>

\begin{abstract}
With phase 1 of the Fast Forward Storage and IO Stack project complete, it is
an excellent opportunity to evaluate many of the decisions made to feed into
the phase 2 effort. The initial effort to define a next generation filesystem
has made admirable contributions in architecture and design. Formalizing the
general idea of data staging as burst buffers for the file system will help
manage the performance variability and offer the data processing opportunities
outside the main compute and file system. Adding a tranasctional mechanism to
manage faults and data visilibilty helps enable effective analytics without
having to work around the IO stack semantics.

@@

This paper examines some of the choices made by the FastForward team and
compares them with other options and offers observations and suggestions based
on these other efforts.  This will include some non-core contributions of other
projects, such as some of the demonstration metadata and data storage
components generated while implementing \DDTns, to make suggestions that may
help the next generation design for how the IO stack works as a whole.

@@

Current production HPC IO stack design is unlikely to offer sufficient features
and performance to adequately serve the needs of an extreme scale platform. To
address these limitations, the US Department of Energy commissioned an effort
to develop a design and prototype for an IO stack suitable for the extreme
scale environment. This is a joint effort led by the Intel Lustre team, Los
Alamos National Laboratory, EMC, and the HDF Group. This team has developed a
specification set~\cite{fastforward:2014:docs} for a future IO stack to address
the identified challenges. The first phase recently completed with a second
phase getting underway. The core focus of the first phase was basic
functionality and design. The second phase will refine this design
incorporating fault recovery and other features missing from the first phase.

@@

The core idea for IOD is to provide a way to manage the IO load that is
separate from the compute nodes and the storage array. Communication intensive
acitivies, such as two-phase, data sieving IO's data rearrangement process, can
be moved to the IOD layer offloading the communication load from the compute
nodes. IOD has three main purposes. First, if the optional burst buffer is
availble, it works as a fast cache absorbing write write operations for the
slower trickle out to the central storage array. It can also be used to
retrieve objects from the central storage array for more efficient read
operations and offer data filtering to make client reads more efficient.
Second, it offers the transaction mechanism for controlling data set visibility
and to manage faults that should prevent a data set from being used. Third,
data processing operations can be placed in the IOD. This operations are
inteded to offer data rearrangement, filtering, and similar operations prior to
it reaching the central storage array.

Offloading the collective two-phase data sieving from the compute nodes to
reorganize data has proven effective at reducing the total time for writing
data due to fewer participants involved in the communication
patterns~\cite{lofstead:2011:nessie-staging}.  Beyond these broad items, there
are many important details some of which are examined in more detail below.

The bottom two layers are the Data Access Object Storage (DAOS) and Versioning
Object Storage Device (VOSD). DAOS is the typical interface that represents
the individual objects stored including the semantics related to ``files''.
VSOD is the actual storage mechanism for the objects defined at the DAOS layer.

The focus of this paper is primarily the IOD layer given the critical role it
has in the performance and functionality of the entire stack. Most of the key
features explored in this paper all have a strong presence in the IOD layer
motivating the focus of this examination.

%Incorporated into the design of these layers and the function of some of these
%layers themsleves, such as the IOD layer, are features to manage both burst IO
%performance and faults. In particular, the IOD layer's use of burst buffers
%attempts to absorb a massive data dump from the compute area that is then
%trickled to the persistent storage area only if desired. Otherwise, it can be
%stored locally on optional SSDs for use by analysis routines eliminating the
%need to use slower rotational media as the intermediary between a simulation
%and the analysis code.

%A second feature is the incorporation of a transaction like mechanism into the
%IOD layer and a related epochs feature into the DAOS layer. This feature offers
%a mechanism to avoid making incomplete data visible to other users and offer a
%level of protection if the data can be stored on the optional SSDs. The
%implications of how these concepts are designed are important to consider for
%the final design. Because they control data visibility and manage faults for
%output operations, they are key features.

The rest of the paper is organized as follows. Section~\ref{sec:burst}
discusses some of the features of incoporating burst buffers as designed and
suggests some considerations and alternatives for the next generation of this
project. Section~\ref{sec:transactions} discusses the transactions approach
offered in the IOD layer and the corresponding epochs in the DAOS layer. It
also offers a comparison to the \DDT system given the very similar high-level
design and motivating use case.  Section~\ref{sec:optimizations} discusses some
of the optimizations offered by the current design and how and when these
optimizations will work best. It also explores some of the dark corners that
may yield performance issues with the current design. 
Section~\ref{sec:summary} disucsses the system overall with recommendations on
what design elements should be considered based on broader issues with current
HPC data centers. The paper is conclued in Section~\ref{sec:conclusion} with
a summary of the broad issues covered in the paper.

@@

By concentrating this functionality into the storage stack, three problems
arise.  First, the amount of network bandwidth, IO bandwidth, and compute power
consumed for example operations from a single application is likely to
completely monopolize the IOD processes. Second, if space and time partitioning
is used instead, the functionality risks being too small to be useful. Third,
the hardware performance advantage for SSDs is questionable. Current NAND-based
flash devices top out at around 400 MB/sec. The key spec that is missing from
this number is that 400 MB/sec is a measure of the fixed number of available
IOPS multipled by the block size. This represents the ideal streaming
performance possible. The problem is that it costs an IOP to read 1 byte or 1
block (4KB or 8KB, depending on the device). It costs 1 IOP to write a full
block--usually. In some cases, it will cost 2 IOPs.  This accounts for the
required pre-erase write prior to writing to a reused block. In the worst case,
it can be 3 IOPS per write. This would be for a partial block write (read the
old block, erase, write the modified block). One-third of 400 MB/sec, about 133
MB/sec, is well below the streaming performance of HDDs.  Granted, there is
still rotational and seek latency to deal with for HDDs, but the advantage for
SSDs has evaporated and potentially turned into a penalty at a considerable
cost premium.  There are faster SSD solutions on the market that incorporate
DRAM for caching and using the PCIe bus, for example, but their price precludes
them from use in an extreme scale platform.

Given these features, the optionality and even incorporation of burst buffers
in the current design should be carefully considered. Much of the advanced, key
functionality proposed as they are currently designed ultimatey relies on the
existence of burst buffers to work. Further thought about how to have an IOD
layer both with and without a burst buffer is required before they can be
considered optional. As the design stands today, they are a required part of
the IOD layer for proper functioning. Unfortunately, it is not clear that they
can address the performance concerns they are intended to cover.

@@

The transaction mechanism manifests in two forms. At the IOD layer, they are
called transactions and are used to judge whether or not an output is complete
or not and control access through treating the transaction ID as a version
identifier. At the DAOS layer, they are called epochs and represent persisted
(durable) transactions from the IOD layer. Each of these offers different
functionality, but are connected as is explained below. How these differ from
the \DDT approach is also explored. While IOD's and \DDTns's transactions are
seemingly very different, they use a similar high-level design, but very
different implementation, to solve the same problem.

@@

\subsection{DAOS Epochs}
The Epoch mechanism differs from transactions. Instead of focusing on when a
particular output is complete, an epoch represents incremental persisted copies
of a container. To simplify the mapping between an IOD transaction and the DAOS
epochs, when an IOD transaction is persisted to DAOS, the IOD transaction ID is
the used as the epoch ID. The key difference is that at the DAOS layer, some
transaction (epoch) IDs will not be represented.

@@

While a pure no metadata model would be ideal for performance, this current
proposal has a few challenges. By providing this basic functionality, metadata
management is split between the user layer and the IOD layer. Consdier the
following items. First, there are no list of containers. This is closer to the
no metadata ideal, but given the rest of the functionality, it exports to the
user level a requirement to track what containers are in the file system.
Second, given the flattening that occurs between IOD and DAOS, the mapping from
IOD objects to DAOS objects is unlikely to be 1-to-1. This is likely a list
instead. Further, given the ability to re-stripe in the IOD layer, the list is
not a list of objects completely included in the IOD object, but instead
objects from which part of the data came from some listed DAOS object.  Third,
sharding IOD objects makes sense based on the compute process connection a
shard has with the complete object.  Striping of an IOD object makes sense in
connection with the DAOS object(s) underneath, but striping within the IOD
layer is not discussed. Further, using striping from clients into the IOD layer
introduces a level of coordination that IOD seeks to avoid.  Fourth, given the
previous items, there is ambiguity with the maximum valid offset of the object.
This could be for the shard or for the collection of shard objects that
represent a distributed variable. Ideally, either a robust metadata management
system should be incorporated into the system, preferably separate from the IOD
layer or all of this functionality should be left to the user layer. The only
potential challenge is determining the container for a file open command and
what the transaction ID should be. This limited operation should be a
straightforward search using a hash or similar fast mechanism for examining a
list of container names and/or object IDs.

@@

NEED TO ENUMERATE SOME OF THE CLIENT-SIDE ONLY PROBLEMS

@@

First, \DDT has a timeout mechanism to detect failures and offer an ability to
recover and clean up the incomplete transactions. IOD's transactions are
asynchronous precluding a relatively simple, short timeout value to detect
failures. Unfortunately, at this point, there is no mechanism in IOD's
transaction handling to detect a fault and clean up any incomplete operations.
While this is intended for phase 2, some initial suggestion on how this might
be handled would give some confidence that the asynchronous approach is viable.

@@

Third, in the single leader model, if the process that manages the transaction
were to fail, there is no ability to abort that transaction. The data will not
be made available because the transaction is still listed as in progress. The
problem is that without a way to clean up this failed transaction, no newer
versions of the same container can be made readable. In reality, the single
leader model would ideally outsource the transaction processing to something
like \DDT on the client side. This does not address the failure case though.

@@

\DDT has addressed this count issue in a couple of ways. First, the
sub-coordinators each have a list of processes from which they expect messages.
Should a message be missed, it is noticed and corrective action can be taken.
Second, \DDT has the concept of sub-transactions. The messaging requirements
are illustrated in Figure~\ref{fig:optimized-protocol}.  Sub-transactions
represent finer grained operations than the entire output, \DDT can manage
multiple writes per client by using a sub-transaction to represent the output
for any item to the file (container). Because of how the sub-transactions are
managed, the singleton sub-transactions, ones in which only a single process
participates, must be declared before the transaction begins so that its
existence can be broadcast as part of the begin transaction message. This
ensures there is global knowledge that the sub-transaction is expected.  That
way if the coordinator (transaction leader) fails, which ever process takes
over that role knows to expect a completion message for that sub-transaction or
the overall transaction cannot complete. Global sub-transactions can be defined
at any time since they are a global, synchronized operation broadcasting their
existence. While this additional layer does introduce messaging, the overhead
is quite small.

The advantages of eliminating these messages is not performance as demonstrated
by the performance of \DDTns. Instead, it offers a much less synchronous model
that matches with different programming models, such as Charm++ or other
task-based approaches. Since it can work for the bulk-synchronous model also,
it is a more broadly applicable approach. This assumes that the observed
potential issues can be addressed successfully.

@@

@@

THiS SECTION NEEDS SOME HELP

@@

\section{Conclusions}
\label{sec:conclusion}

@@

The partial metadata management incorporated into the IOD layer and the lack of
consideration for how to handle and recover from failures are oversights in the
current documents. It is our undertanding that these will be addressed in the
next phase and we hope to help inform that effort with our experiences.

@@
